---
title: "Model fitting"
author: "Meizhen"
date: "`r Sys.Date()`"
abstract: |
  Examine pairwise associations between exposures and whole protein biomarkers using linear regression
  1. Adjusted for age, BMIAdjust race;
  2. Applied inverse probability weighting to account for the nested case–control sampling;
  3. Calculated FDR-adjusted p-values for multiple testing correction.
format: 
  html:
    self-contained: true  # Ensures a single HTML file is created
    code-fold: true
    toc: true
    toc-location: left
theme: united
number-sections: true
editor: visual
fig-width: 18
fig-height: 12
execute: 
  warning: false
---

```{r, warning=FALSE, message=FALSE}

library(tidyverse)
library(ggpubr)
library(data.table)
library(readr)
library(readxl)
library(formattable)
library(OlinkAnalyze)
library(gtsummary)
library(flextable)
library(ggplotify)
library(pheatmap)
library(plotly)
library(gapminder)
library(manipulateWidget)
library(ggfortify)
library(reactable)
library(mice)
library(naniar)
library(ggrepel)
library(writexl)
library(limma)
library(qqman)
library(clusterProfiler)
library(enrichplot)
library(ReactomePA)
library(fgsea)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(biomaRt)
library(mixOmics)
library(sva)

# setwd("/home/meizhen/proteomics/data")

setwd("/Users/meizhenyao/Documents/projects/proteomics_ht/data")

## import data (clean data - remove controls in sample and assay)
data <- fread("intermediate/whole_scale_data.txt")

## Proteins with <LOD% > 75%
OlinkID_lod_remove75<- fread("intermediate/protein_lod_remove75.txt")$OlinkID


chemical_name <- names(data %>% 
                         dplyr::select(starts_with("pfas_"), starts_with("pth_") , starts_with("pah_") , starts_with("parab_") , starts_with("bis_") , starts_with("metal_") , -ends_with("_blod") ) )


data_ready<-  data %>% 
              mutate(ptb_type = case_when(ptb == 0 ~ "Not preterm birth",
                                          ptb == 1 & ptb_spont == 1 ~ "Spontaneous preterm birth",
                                          ptb == 1 & ptb_placental == 1 ~ "Placental preterm birth",
                                          ptb == 1 & ptb_protocol == 1 ~ "Protocol preterm birth"),
                     ptb_cat = factor(ptb, levels = c(0,1), labels = c("No", "Yes")),
                     hispanic = factor(hispanic, levels = c(0,1), labels = c("No", "Yes")),
                     race_recat = factor(ifelse(race == "Caucasian", "Caucasian", "Non-Caucasian"), levels = c("Caucasian", "Non-Caucasian")),
                     parity_recat = factor(case_when(parity == 0 ~ "0",
                                                     parity == 1 ~ "1",
                                                     .default = ">1"), levels = c("0", "1", ">1")),
                     mat_edu_recat = case_when(mat_edu_cat == 1 | mat_edu_cat == 2 ~ 1,
                                               mat_edu_cat == 3 ~ 2,
                                               .default = NULL)) %>% 
              mutate(across(chemical_name, ~(log2(.x)), .names = '{col}_l'),
                     weight = ifelse(ptb == 1, 1/(43/143), 1/(43/1038)),
                     weight_scale = weight/mean(weight))



```

Without weights, but add preterm birth indicator as covariates.

# With IPW

## DEHP

```{r, warning=FALSE, message=FALSE}
protein<- setdiff(names(data_ready %>% dplyr::select(starts_with("OID"))),
                  OlinkID_lod_remove75)

expression<- t(data_ready %>% dplyr::select(protein))

exposure<- names(data_ready %>% 
                   dplyr::select(starts_with("pth") & ends_with("_SG_gm_l")))
```

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
dehp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
dehp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
dehp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "DEHP"

DEHP_gm_weight_sva<- ss.hits.pv

# write.csv(MECPP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MECPP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MECPP

```{r, warning=FALSE, message=FALSE}
protein<- setdiff(names(data_ready %>% dplyr::select(starts_with("OID"))),
                  OlinkID_lod_remove75)

expression<- t(data_ready %>% dplyr::select(protein))

exposure<- names(data_ready %>% 
                   dplyr::select(starts_with("pth") & ends_with("_SG_gm_l")))
```

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mecpp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat+ components$PC1 +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mecpp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mecpp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MECPP"

MECPP_gm_weight_sva<- ss.hits.pv

# write.csv(MECPP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MECPP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MEOHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
meohp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
meohp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
meohp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEOHP"

MEOHP_gm_weight_sva<- ss.hits.pv

# write.csv(MEOHP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEOHP_gm_weight_sva_prolod.csv')
```

Adjust surrogate variable 1

## MEHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mehp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEHP"

MEHP_gm_weight_sva<- ss.hits.pv

# write.csv(MEHP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEHP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MEHHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehhp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehhp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mehhp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEHHP"

MEHHP_gm_weight_sva<- ss.hits.pv

# write.csv(MEHHP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEHHP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MBZP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbzp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbzp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mbzp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MBzP"

MBzP_gm_weight_sva<- ss.hits.pv

# write.csv(MBzP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MBzP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MIBP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mibp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mibp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mibp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MiBP"

MiBP_gm_weight_sva<- ss.hits.pv

# write.csv(MiBP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MiBP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MBP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mbp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MBP"

MBP_gm_weight_sva<- ss.hits.pv

# write.csv(MBP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MBP_gm_weight_sva_prolod.csv')

```

Adjust surrogate variable 1

## MEP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mep_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mep_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mep_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEP"

MEP_gm_weight_sva<- ss.hits.pv

# write.csv(MEP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEP_gm_weight_sva_prolod.csv')
```

Adjust surrogate variable 1

## MCPP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mcpp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mcpp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat +data_ready$ptb_cat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mcpp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MCPP"

MCPP_gm_weight_sva<- ss.hits.pv

# write.csv(MCPP_gm_weight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MCPP_gm_weight_sva_prolod.csv')
```

Adjust surrogate variable 1

## Summary

```{r, warning=FALSE, message=FALSE}
lambda_value<- c(mecpp_value1, mecpp_value2, mecpp_value3,
                 meohp_value1, meohp_value2, meohp_value3,
                 mehp_value1, mehp_value2, mehp_value3,
                 mehhp_value1, mehhp_value2, mehhp_value3,
                 mbzp_value1, mbzp_value2, mbzp_value3,
                 mibp_value1, mibp_value2, mibp_value3,
                 mbp_value1, mbp_value2, mbp_value3,
                 mep_value1, mep_value2, mep_value3,
                 mcpp_value1, mcpp_value2, mcpp_value3)

lambda_model<- c('mecpp_value1', 'mecpp_value2', 'mecpp_value3',
                 'meohp_value1', 'meohp_value2', 'meohp_value3',
                 'mehp_value1', 'mehp_value2', 'mehp_value3',
                 'mehhp_value1', 'mehhp_value2', 'mehhp_value3',
                 'mbzp_value1', 'mbzp_value2', 'mbzp_value3',
                 'mibp_value1', 'mibp_value2', 'mibp_value3',
                 'mbp_value1', 'mbp_value2', 'mbp_value3',
                 'mep_value1', 'mep_value2', 'mep_value3',
                 'mcpp_value1', 'mcpp_value2', 'mcpp_value3')
sva_value<- data.frame(lambda_model, 
                       lambda_value) %>% 
            mutate(exposure = factor(sub("_.*", "", lambda_model),
                                     levels = c('mecpp','meohp', 'mehp','mehhp','mbzp','mibp','mbp','mep','mcpp')),
                   model = sub(".*_", "", lambda_model))


sva_plot <- ggplot(sva_value, aes(x = model, y = lambda_value, color = exposure, group = exposure)) +
  geom_point(size = 6) +
  geom_line(size = 2) +
  geom_text(aes(label = round(lambda_value, 2)), vjust = -1, size = 10) +
  facet_wrap(~ exposure, scales = "free_x") +
  labs(x = "Model", y = "Lambda Value", title = "Lambda Values by Model for Each Phthalate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        text = element_text(size = 20),
        legend.position = "none")

sva_plot
```

Model 1: Model 2: Adjust PC1 Model 3: Adjust surrogate variable 1

# Without IPW

## DEHP

```{r, warning=FALSE, message=FALSE}
protein<- setdiff(names(data_ready %>% dplyr::select(starts_with("OID"))),
                  OlinkID_lod_remove75)

expression<- t(data_ready %>% dplyr::select(protein))

exposure<- names(data_ready %>% 
                   dplyr::select(starts_with("pth") & ends_with("_SG_gm_l")))
```

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$mat_edu_recat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("λ = ", round(lambda,3)))
dehp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")


sum(ss.hits.pv$adj.P.Val<0.2) 

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("λ = ", round(lambda,3)))
dehp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


mod <- model.matrix(~ data_ready$pth_dehp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), adjust.method = "BY",sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("λ = ", round(lambda,3)))

sum(ss.hits.pv$P.Val<0.05) 
dehp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "DEHP"

DEHP_gm_noweight_sva<- ss.hits.pv

# write.csv(DEHP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/DEHP_gm_noweight_sva_prolod_ad.csv')
# 
```

Adjust surrogate variable 1

## MECPP

```{r, warning=FALSE, message=FALSE}
protein<- setdiff(names(data_ready %>% dplyr::select(starts_with("OID"))),
                  OlinkID_lod_remove75)

expression<- t(data_ready %>% dplyr::select(protein))

exposure<- names(data_ready %>% 
                   dplyr::select(starts_with("pth") & ends_with("_SG_gm_l")))
```

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mecpp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")


sum(ss.hits.pv$adj.P.Val<0.2) 

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mecpp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), adjust.method = "BH",sort.by = "none") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

sum(ss.hits.pv$adj.P.Val<0.2) 
mecpp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MECPP"

MECPP_gm_noweight_sva<- ss.hits.pv

write.csv(MECPP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MECPP_gm_noweight_sva_prolod_ad.csv')

```

Adjust surrogate variable 1

## MEOHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
meohp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
meohp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

sum(ss.hits.pv$adj.P.Val<0.2) 

sum(ss.hits.pv$P.Value<0.05) 
meohp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEOHP"

MEOHP_gm_noweight_sva<- ss.hits.pv

write.csv(MEOHP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEOHP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MEHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mehp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEHP"

MEHP_gm_noweight_sva<- ss.hits.pv

write.csv(MEHP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEHP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MEHHP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehhp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mehhp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat  )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mehhp_value3<- lambda

ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEHHP"

MEHHP_gm_noweight_sva<- ss.hits.pv

write.csv(MEHHP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEHHP_gm_noweight_sva_prolod_ad.csv')

```

Adjust surrogate variable 1

## MBZP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbzp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbzp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat  )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mbzp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MBzP"

MBzP_gm_noweight_sva<- ss.hits.pv

write.csv(MBzP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MBzP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MIBP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mibp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mibp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mibp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MiBP"

MiBP_gm_noweight_sva<- ss.hits.pv

write.csv(MiBP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MiBP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MBP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mbp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}

# summary(lm(pth_mbp_SG_gm_l ~ age + bmi_prepreg + race_recat, data = data_ready))

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg+ data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mbp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MBP"

MBP_gm_noweight_sva<- ss.hits.pv

write.csv(MBP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MBP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MEP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mep_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mep_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")#, sort.by = "none"



lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mep_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MEP"

MEP_gm_noweight_sva<- ss.hits.pv

write.csv(MEP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MEP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## MCPP

### Model 1

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mcpp_value1<- lambda


```

### Model 2

```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(protein), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)

weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l+ components$PC1+ data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
out <- lmFit(expression,mod)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH")

lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
mcpp_value2<- lambda


```

Adjust PC1

### Model 3

```{r, warning=FALSE, message=FALSE}


weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat )
mod0 = model.matrix(~1,data=data_ready)

n.sv = num.sv(expression,mod,method="leek")
svobj = sva(expression,mod,mod0,n.sv=1)

modSv = cbind(mod,svobj$sv)

out <- lmFit(expression,modSv)
out <- eBayes(out, trend = TRUE, robust = TRUE)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none", adjust.method = "BH") #, sort.by = "none"


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


sum(ss.hits.pv$adj.P.Val<0.2) 
sum(ss.hits.pv$P.Value<0.05) 
mcpp_value3<- lambda


ss.hits.pv$Outcome<- rownames(ss.hits.pv)
ss.hits.pv$Exposure<- "MCPP"

MCPP_gm_noweight_sva<- ss.hits.pv

write.csv(MCPP_gm_noweight_sva, '/Users/meizhenyao/Documents/projects/proteomics_ht/data/model/limma/MCPP_gm_noweight_sva_prolod_ad.csv')
```

Adjust surrogate variable 1

## Summary

```{r, warning=FALSE, message=FALSE}
# lambda_value<- c(dehp_value1, dehp_value2, dehp_value3,
#                  mecpp_value1, mecpp_value2, mecpp_value3,
#                  meohp_value1, meohp_value2, meohp_value3,
#                  mehp_value1, mehp_value2, mehp_value3,
#                  mehhp_value1, mehhp_value2, mehhp_value3,
#                  mbzp_value1, mbzp_value2, mbzp_value3,
#                  mibp_value1, mibp_value2, mibp_value3,
#                  mbp_value1, mbp_value2, mbp_value3,
#                  mep_value1, mep_value2, mep_value3,
#                  mcpp_value1, mcpp_value2, mcpp_value3)
# 
# lambda_model<- c('dehp_value1', 'dehp_value2', 'dehp_value3',
#                  'mecpp_value1', 'mecpp_value2', 'mecpp_value3',
#                  'meohp_value1', 'meohp_value2', 'meohp_value3',
#                  'mehp_value1', 'mehp_value2', 'mehp_value3',
#                  'mehhp_value1', 'mehhp_value2', 'mehhp_value3',
#                  'mbzp_value1', 'mbzp_value2', 'mbzp_value3',
#                  'mibp_value1', 'mibp_value2', 'mibp_value3',
#                  'mbp_value1', 'mbp_value2', 'mbp_value3',
#                  'mep_value1', 'mep_value2', 'mep_value3',
#                  'mcpp_value1', 'mcpp_value2', 'mcpp_value3')

lambda_value<- c(dehp_value1, dehp_value3,
                 mecpp_value1, mecpp_value3,
                 meohp_value1, meohp_value3,
                 mehp_value1, mehp_value3,
                 mehhp_value1, mehhp_value3,
                 mbzp_value1, mbzp_value3,
                 mibp_value1, mibp_value3,
                 mbp_value1, mbp_value3,
                 mep_value1, mep_value3,
                 mcpp_value1, mcpp_value3)

lambda_model<- c('dehp_value1', 'dehp_value2',
                 'mecpp_value1', 'mecpp_value2',
                 'meohp_value1', 'meohp_value2', 
                 'mehp_value1', 'mehp_value2', 
                 'mehhp_value1', 'mehhp_value2',
                 'mbzp_value1', 'mbzp_value2', 
                 'mibp_value1', 'mibp_value2', 
                 'mbp_value1', 'mbp_value2',
                 'mep_value1', 'mep_value2', 
                 'mcpp_value1', 'mcpp_value2')

sva_value<- data.frame(lambda_model, 
                       lambda_value) %>% 
            mutate(exposure = factor(toupper(sub("_.*", "", lambda_model)),
                                     levels = toupper(c('dehp','mecpp','meohp', 'mehp','mehhp','mbzp','mibp','mbp','mep','mcpp'))),
                   model = sub(".*_", "", lambda_model))


sva_plot <- ggplot(sva_value, aes(x = model, y = lambda_value, color = exposure, group = exposure)) +
  geom_point(size = 6) +
  geom_line(size = 2) +
  geom_text(aes(label = round(lambda_value, 2)), vjust = -1, size = 10) +
  ylim(0, 8) +
  facet_wrap(~ exposure, scales = "free_x") +
  labs(x = "Model", y = "Lambda Value", title = "Lambda Values by Model for Each Phthalate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 20),
        axis.title = element_text(colour = "black", face = "bold", size = 18),
        text = element_text(size = 20),
        legend.position = "none",
        strip.text = element_text(colour = "black", face = "bold", size = 18))

sva_plot
```

Model 1: Model 2: Adjust PC1 Model 3: Adjust surrogate variable 1
