---
title: "Model fitting"
author: "Meizhen"
date: "`r Sys.Date()`"
abstract: |
  Examine pairwise associations between exposures and whole protein biomarkers using linear regression, and 
  1. Adjusted for age, BMI, and race;
  2. Applied inverse probability weighting to account for the nested case–control sampling;
  3. Calculated FDR-adjusted p-values for multiple testing correction.
format: 
  html:
    self-contained: true  # Ensures a single HTML file is created
    code-fold: true
    toc: true
    toc-location: left
theme: united
number-sections: true
editor: visual
fig-width: 18
fig-height: 12
execute: 
  warning: false
---

```{r, warning=FALSE, message=FALSE}

library(tidyverse)
library(ggpubr)
library(data.table)
library(readr)
library(readxl)
library(formattable)
library(OlinkAnalyze)
library(gtsummary)
library(flextable)
library(ggplotify)
library(pheatmap)
library(plotly)
library(gapminder)
library(manipulateWidget)
library(ggfortify)
library(reactable)
library(mice)
library(naniar)
library(ggrepel)
library(writexl)
library(limma)
library(qqman)
library(clusterProfiler)
library(enrichplot)
library(ReactomePA)
library(fgsea)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(biomaRt)
library(mixOmics)

# setwd("/home/meizhen/proteomics/data")

setwd("/Users/meizhenyao/Documents/projects/proteomics_ht/data")

## import data (clean data - remove controls in sample and assay)
data <- fread("intermediate/whole_scale_data.txt")

chemical_name <- names(data %>% 
                         dplyr::select(starts_with("pfas_"), starts_with("pth_") , starts_with("pah_") , starts_with("parab_") , starts_with("bis_") , starts_with("metal_") , -ends_with("_blod") ) )


data_ready<-  data %>% 
              mutate(ptb_type = case_when(ptb == 0 ~ "Not preterm birth",
                                          ptb == 1 & ptb_spont == 1 ~ "Spontaneous preterm birth",
                                          ptb == 1 & ptb_placental == 1 ~ "Placental preterm birth",
                                          ptb == 1 & ptb_protocol == 1 ~ "Protocol preterm birth"),
                     ptb_cat = factor(ptb, levels = c(0,1), labels = c("No", "Yes")),
                     hispanic = factor(hispanic, levels = c(0,1), labels = c("No", "Yes")),
                     race_recat = factor(ifelse(race == "Caucasian", "Caucasian", "Non-Caucasian"), levels = c("Caucasian", "Non-Caucasian")),
                     parity_recat = factor(case_when(parity == 0 ~ "0",
                                                     parity == 1 ~ "1",
                                                     .default = ">1"), levels = c("0", "1", ">1"))) %>% 
              mutate(across(chemical_name, ~(log2(.x)), .names = '{col}_l'),
                     weight = ifelse(ptb == 1, 1/(43/143), 1/(43/1038)),
                     weight_scale = weight/mean(weight))



```


# MECPP
```{r, warning=FALSE, message=FALSE}
protein<- names(data_ready %>%
                  dplyr::select(starts_with("OID")))

expression<- t(data_ready %>% dplyr::select(starts_with("OID")))

exposure<- names(data_ready %>% 
                   dplyr::select(starts_with("pth") & ends_with("_SG_gm_l")))
```

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$ptb_cat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```



## with weight, but add pca1 of protein as covariates

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mecpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


# MEOHP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_meohp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```



# MEHP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mehp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


# MEHHP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mehhp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```




# MBZP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mbzp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```















# MIBP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mibp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


# MBP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mbp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


# MEP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mep_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


# MCPP

## with weight

```{r, warning=FALSE, message=FALSE}
weight_matrix<- matrix(rep(data_ready$weight, nrow(expression)),
                             nrow = nrow(expression),
                             byrow = TRUE)

mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression,mod,weights=weight_matrix)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

weight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


## without weight

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat)
out <- lmFit(expression, mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready), sort.by = "none")

noweight<- ss.hits.pv$P.Value
lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```


```{r, warning=FALSE, message=FALSE}
## compare
plot(-log10(noweight), -log10(weight),
     xlab = "Unweighted p-values",
     ylab = "Weighted p-values",
     main = "Comparison of P-values: Weighted vs Unweighted",
     pch = 16, col = "#1f77b4") 

abline(a = 0, b = 1, col = "red", lty = 2) 

sum(noweight<0.05) 
sum(weight<0.05) 


# lead to an overestimation of the number of significant associations or genes. Your weights (e.g., 143/43 vs. 1038/43) are inverse sampling probabilities. These are not appropriate as-is for limma, which expects weights to reflect heteroscedasticity, i.e., more weight for more precise samples.
```


## without weight, but add preterm birth as covariates???

```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$ptb_cat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


```


## without weight, but add parity as covariates
```{r, warning=FALSE, message=FALSE}
mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat)
fit <- lmFit(expression,mod)
out <- eBayes(fit)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))


# 1. Calculate the fitted values
fitted_values <- fit$coefficients %*% t(mod)

# 2. Calculate the residuals: observed - fitted
residuals <- expression - fitted_values
```

## without weight, but add pca1 of residual as covariates

```{r, warning=FALSE, message=FALSE}
pc <- pca(t(residuals), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))

```

## without weight, but add pca1 of protein as covariates
```{r, warning=FALSE, message=FALSE}
pc <- pca(data_ready %>% dplyr::select(starts_with("OID")), ncomp = 10)
pc$cum.var[10]
components<- data.frame(pc$variates$X)


mod <- model.matrix(~ data_ready$pth_mcpp_SG_gm_l + data_ready$age + data_ready$bmi_prepreg + data_ready$race_recat + data_ready$parity_recat+ components$PC1)
out <- lmFit(expression,mod)
out <- eBayes(out)
ss.hits.pv <- topTable(out, coef=2, number=ncol(data_ready))


lambda <- qchisq(median(ss.hits.pv$P.Value, na.rm=T), df = 1, lower.tail = F) / qchisq(0.5, 1)
qqman::qq(ss.hits.pv$P.Value, main=paste0("Lambda = ", round(lambda,3)))
```












